Statistical mechanics of Wigner negativity and magic states in quantum computation

### 2021-06-07 magic_paper.tex
Trialing a new title.
Potentially better title.
Changes made, some rearrangement required

\begin{document}
% Labels and Refs (consistency, cref vs ref, thm vs lem vs prop, ??)
% Figs check(fig.0, move to app, consistency)

% Notation consistency (R is rate not bound, noise vs error, \ln{} -> \ln, italics, keep \bmx for vectors?)
% UK vs US english
% Spacing (figs, sections, paragraphs, eq. numbering, \noindent, QEDs)

@@

\title{General constraints on magic distillation protocols from statistical mechanics}

@@

\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Magic distillation protocols \& stabilizer reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with noise parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the noise parameters $\epsilon, \epsilon'$, and the free energy the protocol channel $\E$ injects in. In addition a quantity $\phi$ links the computational and physical aspects, and corresponds to the degree to which Wigner negativity in $\rho$ can be assigned a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}

@@

The phase-point operators form an orthogonal operator basis with respect to the Hilbert-Schmidt inner product, as discussed in~\cref{app:wigner}.
Therefore, any quantum state $\rho \in \cal{B}(\cal{H}_d)$ can be expressed as a linear combination of them,
\begin{equation}
    \rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz},
\end{equation}
where the coefficient vector $\W{\rho}$ is the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmx]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmx} \rho].
\end{equation}

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmx)$ is readily seen to be a $d^2$-dimensional quasi-probability distribution over $\cal{P}_d$. More precisely, $W_\rho(\x)$ is a bounded real-valued function on $\P_d$ with the normalisation property $\sum_{\x} W_\rho(\bmx) = 1$ (see~\cref{app:wigner} for details).  In~\cref{fig:wstate_examples}, we show Wigner distributions of different types of qutrit states.

\begin{figure}[t]
    \centering
    \subfigure[][]{%
    \label{fig:maxmix}%
    \includegraphics[height=2cm]{figs/maxmixed.pdf}
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:zero}%
    \includegraphics[height=2cm]{figs/zerostate.pdf}
    }\\
    \subfigure[][]{%
    \label{fig:bound}%
    \includegraphics[height=2cm]{figs/boundstate.pdf}
    }\hspace{8pt}%
    \subfigure[][]{%
    \label{fig:Strange}%
    \includegraphics[height=2cm]{figs/Strangestate.pdf}
    }
    \caption{\textbf{Qutrit Wigner distributions of varying magic.} 
    \subref{fig:maxmix} Maximally mixed state $\frac{1}{3}\id$; \subref{fig:zero} Stabilizer zero state $|0\>\<0|$; \subref{fig:bound} A non-stabiliser Wigner-positive state; \subref{fig:Strange} Magic Strange state $\ket{{\rm{S}}} = \frac{1}{\sqrt{2}}(\ket{1} - \ket{2})$, coined in~\cite{cit:veitch2}.
    }%
    \label{fig:wstate_examples}
\end{figure}

@@

\begin{lemma} Let $\w \in \mathbb{R}^n,\w'\in \mathbb{R}^m$ be quasi-distributions, and let $\r \in \mathbb{R}^n, \r' \in \mathbb{R}^m$ be probability distributions with non-zero components. Then $(\w, \r) \succ (\w', \r')$ if and only if $L_{\w|\r}(x) \ge L_{\w |\r}(x)$ for all $x \in [0,1]$.
\end{lemma}

@@

\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ into $\sigma$--fragments.} 
	Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) are labelled, with other established magic theories also contained within the yellow regions.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$. 
    Each $\O_\sigma$ is extensible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment, relative majorisation can be used allowing for a tractable approach towards the study of magic state interconversions.
    }
    \label{fig:zoo}
\end{figure}

@@

This figure includes numerical estimates of the optimal majorization bounds $R_{\rm{opt}}$ obtained by considering the full Lorenz curve, which show that the first elbow constraint can be further improved.
\begin{figure}[t]
    \centering
    \includegraphics[scale=0.35]{figs/distill_bounds.pdf}
    \caption{\textbf{Distillation bounds in the unital fragment.} Distillation bounds obtained by the first elbow majorisation constraint $R$, mana $\M$ and max--thauma $\theta_{\rm{max}}$ are plotted for $\epsilon' = 0$, up to error $\epsilon = 3/7$.
    Majorisation provides stricter rates than the global mana and max--thauma bounds.
    Numerical bounds $R_{\rm{opt}}$ obtained by full Lorenz curve comparison for some Strange state purification processes with even number of copies $m,n \leq 30$ highlight that the first elbow constraint can be improved.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

However in this direction, we first note that the analysis for the temperature-dependent bound given by Equation (\ref{eq:rate_bounds_proof}) implicitly makes use of the single-shot Renyi divergence $D_\infty (\p ||\r)$, which is given by
\begin{equation}
D_\infty (\p ||\r) = \log \max_i \frac{p_i}{r_i}.
\end{equation}
This does remain well-defined on quasi-distributions, and from the majorization relation also has a clear physical meaning.
This motivates defining the following \emph{Wigner divergence} $D_\infty(\rho, \tau) := D_\infty (W_\rho(\x) || W_\tau (\x))$ on all quantum states. In Appendix~\ref{app:lcst_technical}, we show that this divergence has the following natural properties that are derived via relative majorization.
\begin{theorem} Let $\tau$ be in the interior of $\F$, and define the Wigner divergence $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$.
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
See Appendix~\ref{app:lcst_technical} for the proof of this.

We can now go further, and show that a range of entropies remain both well-defined and meaningful in this setting. The $\alpha$--Renyi entropy $H_\alpha(\p)$ is defined as
\begin{equation}
	H_\alpha(\p) := \frac{1}{1-\alpha} \log \sum_i p_i^\alpha.
\end{equation}
To extend to quasi-distributions we must ensure the following: (a) the entropy is well-defined on quasi-distributions, and (b) obeys monotonicity under stabilizer/free operations. In the Appendices we show the following.
\begin{theorem} If $\alpha = \frac{2a}{2b-1}$ for integers $a,b \in \mathbb{N}$ and $a\ge b$ then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions $\w$, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'$ then $H_\alpha (\w) \le H_\alpha(\w')$.
\end{theorem}
This therefore provides a meaningful extension of entropies to quasi-distributions. Moreover we can prove the following result, which establishes the equivalence between Wigner negativity and the existence of negative Renyi entropy.
\begin{theorem} A quantum state $\rho$ has Wigner negativity if and only if $H_\alpha(W_\rho) <0$ for some $\alpha =  \frac{2a}{2b-1}$, with integers $a \ge b$.
\end{theorem}
Therefore, the statistical mechanical description of stabilizer operations provides a setting in which negative entropy is fully meaningful, and quantifies the fact that magic states are `more ordered' than a perfectly sharp, deterministic classical state $\w = (1,0, \dots , 0)$ with zero entropy. We note that prior work has shown that negative conditional entropy \ddd{[CITE: ``The thermodynamic meaning of
negative entropy'']} arises in the context of quantum correlations, however this is unrelated to the present negative entropy.

The $\alpha \rightarrow 1$ limit of the Renyi entropy is not well-defined, however the Renyi entropy description shows that this diverges to $-\infty$ due to the Lorenz curve $L_\rho(x)$ no longer being bounded by $1$. Alternatively, we can write the divergence of the entropy as $-\epsilon^{-1} \log \sum_{\x} |W_\rho (\x)|^{1+\epsilon}$ with $\epsilon \rightarrow 0^+$, which leads to a second new description of mana: \emph{it is the rate of divergence of the Renyi-entropy} as we approach the limiting Shannon entropy $H(\w)$. We also note the recent observation \ddd{[CITE Renyi Entropy and Free Energy by John C. Baez]}that in standard statistical mechanics that $H_\alpha(\p)$ can be related to the ``$q$--derivative'' of free energy. Therefore, we might speculate if the presence of magic in Clifford circuits admits a rigorous interpretation in terms of a phase transition signalled by a first-order divergence in the Wigner free energy. However since variation in $\alpha$ is no longer smooth ($\alpha$ is restricted to a dense set of rational values) technical obstacles exist and we do not pursue this speculative line here, but just note that other works exist in which computational transitions are considered in relation to phase transitions \ddd{[Need to cite]}.

However, every entropy is best understand as a relative entropy and so we now state the extension to Renyi divergences. \ddd{[Add statements of monotonicity under free ops]}

We can now apply this to obtain new magic bounds. \ddd{[Add]}
\section{Outlook and open questions}
\label{sec:lower_bounds}

@@

\begin{theorem} Let $\tau$ be in the interior of $\F$, and define $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$. \nick{annoying that it doesn't agree with the simpler property of $D_\alpha$ for finite $\alpha$ in~\cref{thm:Da_props}}
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
\begin{proof}
	Since $\tau$ is in the interior of $\F$, its Wigner function obeys $W_\tau(\x) >0$ for all points $\x$ in the phase space. In general, $W_\rho(\x)$ is a quasi-distribution, however since $\sum_{\x} W_\rho (\x) = 1$ this implies it must have at least one positive component, so $\max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)}>0$. Therefore $D_\infty (\rho,\tau)$ is always well-defined.

@@

\begin{table}[h]
  \def\arraystretch{1.5}
  \centering
  \begin{tabular}{c|c|c|c||c|c}
    $w_i$ & $w_{i+1}$ & $\alpha$ & $\frac{w_i^{\alpha-1}}{\alpha-1} \geq \frac{w_{i+1}^{\alpha-1}}{\alpha-1}$ & $\alpha$ & $\frac{w_i^{\alpha-1}}{\alpha-1} \geq \frac{w_{i+1}^{\alpha-1}}{\alpha-1}$ \\[0.5ex]\hline
    $2$ & $1.5$  & $3$ & $2 > 1.125$ T & $1/3$ & $-0.94 > -1.14$ T \\
    $2$ & $-1.5$ & $3$ & $2 > 1.125$ T & $1/3$ & $-0.94 > -1.14$ T \\
    $2$ & $-2.5$ & $3$ & $2 < 3.125$ F & $1/3$ & $-0.94 < -0.81$ F \\[0.5ex]\hline
    $2/3$ & $1/3$  & $3$ & $0.22 > 0.06$ T & $1/3$ & $-1.97 > -3.12$ T \\
    $2/3$ & $-1/3$ & $3$ & $0.22 > 0.06$ T & $1/3$ & $-1.97 > -3.12$ T \\
    $2/3$ & $-4/3$ & $3$ & $0.22 < 0.89$ F & $1/3$ & $-1.97 < -1.24$ F \\[0.5ex]\hline
    $2$ & $1.5$  & $4$ & $2.67 > 1.125$ T & $2/3$ & $-2.38 > -2.62$ T \\
    $2$ & $-1.5$ & $4$ & $2.66 > -1.125$ T & $2/3$ & $-2.38 > 2.62$ F \\
    $2$ & $-2.5$ & $4$ & $2.66 > 5.21$ T & $2/3$ & $-2.38 > 2.21$ F \\
    $-2$ & $-2.5$ & $4$ & $-2.66 > -5.21$ T & $2/3$ & $2.38 > 2.21$ T \\[0.5ex]\hline
    $2/3$ & $1/3$  & $4$ & $0.10 > 0.01$ T & $2/3$ & $-3.43 > -4.33$ T \\
    $2/3$ & $-1/3$ & $4$ & $0.10 > -0.01$ T & $2/3$ & $-3.43 > 4.33$ F \\
    $2/3$ & $-4/3$ & $4$ & $0.10 > -0.79$ T & $2/3$ & $-3.43 > 2.73$ F \\
    $-1/3$ & $-2/3$ & $4$ & $-0.01 > -0.10$ T & $2/3$ & $4.33 > 3.43$ T \\
    $-1/3$ & $-4/3$ & $4$ & $-0.01 > -0.79$ T & $2/3$ & $4.33 > 2.73$ T
  \end{tabular}
  \caption{\nick{NUMERICS ON SCHUR-OSTROWSKI CONDITION} The case $\alpha=$ even$/$odd and $\alpha > 1$ works for any combination of $w_i, w_{i+1}$ with $w_i > w_{i+1}$. 
  All other cases do not work.
  I put it here cause it's clunky}
  \label{tab:test}
\end{table}

@@

\ddd{[ISSUE WITH NOTATION: We can't clash with the quantum Renyi entropies. I think we have to use $H_\alpha(W_\rho)$ is bulky...but hopefully clear.]}

We define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(W_\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(W_\rho)$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(W_\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
Inducing any magic state $\rho$ with depolarising noise, $\rho(\epsilon) = (1-\epsilon)\rho + \epsilon \sigma$ results in $H_{\alpha}\left(\rho(\epsilon)\right) \leq H_\alpha(\rho)$ because there exists a free operation $\E \in \O$ that transforms $\rho(\epsilon) = \E(\rho)$ and $H_\alpha$ is Schur-concave.

For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $H_\alpha$ is decreasing in $\epsilon$ (not Schur-conocave for $\alpha < 1$) and increasing in $\alpha$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:Htest}
\end{figure}

@@

\begin{lemma}
	Let $\alpha \geq \beta > 1$ where $\alpha, \beta$ are even$/$odd.
	Then,
	\begin{equation}
		H_\alpha(\x) \geq H_\beta(\x) \text{ and } D_\alpha(\x || \r) \leq D_\beta(\x || \r)
	\end{equation}
	for any quasi-distribution $\x \in \mathbb{R}^N$ and probability distribution $\r \in \mathbb{R}^N$.
\end{lemma}
\begin{proof}
	For $\alpha \geq \beta > 1$, we have
	\begin{equation}
		0 > \frac{1}{1-\alpha} \geq \frac{1}{1-\beta},
	\end{equation}
	and $w_i^\alpha = |w_i|^\alpha \leq |w_i|^\beta = w_i^\beta$.

	\nick{how to proceed?}

	For any $\r$ with positive rational components, we have $D_\alpha(\x || \r) \leq D_\beta(\x || \r)$ due to~\cref{thm:H2D}.
\end{proof}

We can prove additivity of $H_{\alpha}$ on quantum states.

\begin{lemma}\label{H_add}
	For any quantum states $\rho, \rho'$, we have
	\begin{equation}
		H_{\alpha}(\rho \otimes \rho') = H_{\alpha}(\rho) + H_{\alpha}(\rho'),
	\end{equation}
	where $\hspace{-1pt}\alpha\hspace{-1pt}$ is such that $\hspace{-1pt}H_\alpha(\sigma)\hspace{-2pt}$ is well-defined for all states $\sigma$.
\end{lemma}
\begin{proof}
	This can be seen directly by the multiplicative property of the Wigner distribution,
	\begin{align*}
		H_{\alpha}&(\rho \otimes \rho') = \frac{1}{1-\alpha}\log \sum_{\bmx,\bmx'} \left[W_{\rho}(\bmx)W_{\rho}(\bmx')\right]^{\alpha} \\
		&= \frac{1}{1-\alpha}\left( \log \sum_{\bmx}W_{\rho}(\bmx)^{\alpha} + \log \sum_{\bmx'}W_{\rho}(\bmx')^{\alpha} \right) \\
		&= H_{\alpha}(\rho) + H_{\alpha}(\rho').
	\end{align*}
\end{proof}

@@

Example 1: if $\alpha =$ odd$/$odd, then $f(w)$ is not monotonous because the sign of $f'(w)$ would depend on the sign of $w$. 
In~\cref{tab:test}, this case fails when $w_i > 0 > -w_i > w_{i+1}$.

>>>>

\begin{document}
% Labels and Refs (consistency, cref vs ref, thm vs lem vs prop, ??)
% Figs check(fig.0, move to app, consistency, cbars, iffalse)

% Notation consistency (noise vs error, \ln{} -> \ln, italics, Aberg -> {\AA}berg)
% UK vs US english
% Spacing (figs, sections, paragraphs, eq. numbering, \noindent, QEDs)

@@

\title{Statistical mechanics of Wigner negativity and magic states in quantum computation}

@@

\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/protocol_diagram.pdf}
    \caption{\textbf{Magic distillation protocols \& stabilizer reference states.} 
	A magic protocol (top figure) converts $n$ copies of some noisy magic state $\rho$ with noise parameter $\epsilon$ to $m$ copies of some less noisy state $\rho'$ with $\epsilon' < \epsilon$. The physics of the protocol can be analysed by considering how it would transform some distinguished reference stabilizer state $\tau^{\otimes n}$ (bottom figure). For the case where this reference state is a thermal state at some temperature $T$ our magic distillation bounds are functions of $T$, the noise parameters $\epsilon, \epsilon'$, and the free energy the protocol channel $\E$ injects in. In addition, a quantity $\phi$ links the computational and physical aspects, and corresponds to the degree to which Wigner negativity in $\rho$ can be assigned a sharp energy via the system Hamiltonian.}
    \label{fig:sketch}
\end{figure}

@@

The phase-point operators form an orthogonal operator basis with respect to the Hilbert-Schmidt inner product, as discussed in~\cref{app:wigner}.
Therefore, any quantum state $\rho \in \cal{B}(\cal{H}_d)$ can be expressed as a linear combination of them, $\rho = \sum_{\bmz \in \cal{P}_d} \W[\bmz]{\rho} A_{\bmz}$, where the coefficient vector $\W{\rho}$ is the Wigner distribution of state $\rho$,
\begin{equation}\label{eq:wstate}
    \W[\bmx]{\rho} \coloneqq \frac{1}{d}\tr[A_{\bmx} \rho].
\end{equation}

For any quantum state $\rho$, the Wigner distribution $W_\rho(\bmx)$ is readily seen to be a $d^2$-dimensional quasi-probability distribution over $\cal{P}_d$. More precisely, $W_\rho(\x)$ is a bounded real-valued function on $\P_d$ with the normalisation property $\sum_{\x} W_\rho(\bmx) = 1$ (see~\cref{app:wigner} for details).

@@

\nick{Is the following result more distinguished than other results we have in the appendices?}
\begin{lemma} Let $\w \in \mathbb{R}^n,\w'\in \mathbb{R}^m$ be quasi-distributions, and let $\r \in \mathbb{R}^n, \r' \in \mathbb{R}^m$ be probability distributions with non-zero components. Then $(\w, \r) \succ (\w', \r')$ if and only if $L_{\w|\r}(x) \ge L_{\w |\r}(x)$ for all $x \in [0,1]$.
\end{lemma}

@@

\iffalse
\begin{figure}[t]
    \centering
        \includegraphics[scale=0.3]{figs/operations.pdf}
    \caption{\textbf{Decomposition of a magic theory $\R$ into $\sigma$--fragments.} 
	Examples of magic theories ($\so$: Stabilizer operations, $\Omax$: Completely positive-Wigner-preserving operations, $\rcu$: Random Clifford Unitaries -- subclass of $\so$) are labelled, with other established magic theories also contained within the yellow regions.
    We introduce $\sigma$--fragments $\O_\sigma$ defined for all free states $\sigma$ that cover $\Omax$. 
    Each $\O_\sigma$ is extensible to a set of stochastic maps outside the $\cptp$ operations.
    Within each $\sigma$--fragment, relative majorisation can be used allowing for a tractable approach towards the study of magic state interconversions.
    }
    \label{fig:zoo}
\end{figure}
\fi

@@

This figure includes numerical estimates of the optimal majorization bounds $R_{\rm{opt}}$ obtained by considering the full Lorenz curve, which show that the first elbow constraint can be further improved.
\begin{figure}[t]
    \centering
    \includegraphics[scale=0.35]{figs/distill_bounds.pdf}
    \caption{\textbf{Distillation bounds in the unital fragment.} Distillation bounds obtained by the first elbow majorisation constraint $R$, mana $\M$ and max--thauma $\theta_{\rm{max}}$ are plotted for $\epsilon' = 0$, up to error $\epsilon = 3/7$.
    Majorisation provides stricter rates than the global mana and max--thauma bounds.
    Numerical bounds $R_{\rm{num}}$ obtained by full Lorenz curve comparison for some Strange state purification processes with even number of copies $m,n \leq 30$ highlight that the first elbow constraint can be improved.
    }
    \label{fig:distill_bounds}
\end{figure}

@@

However in this direction, we first note that the analysis for the temperature-dependent bound given by Equation (\ref{eq:rate_bounds_proof}) implicitly makes use of the single-shot Renyi divergence $D_\infty (\p ||\r)$, which is given by
\begin{equation}
D_\infty (\p ||\r) = \log \max_i \frac{p_i}{r_i}.
\end{equation}
This does remain well-defined on quasi-distributions, and from the majorization relation also has a clear physical meaning.
This motivates defining the following \emph{Wigner divergence} $D_\infty(\rho, \tau) := D_\infty (W_\rho(\x) || W_\tau (\x))$ on all quantum states. In~\cref{app:lcst_technical}, we show that this divergence has the following natural properties that are derived via relative majorization.
\begin{theorem} Let $\tau$ be in the interior of $\F$, and define the Wigner divergence $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$.
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
See Appendix~\ref{app:lcst_technical} for the proof of this.

We can now go further, and show that a range of entropies remain both well-defined and meaningful in this setting. The $\alpha$--Renyi entropy $H_\alpha(\p)$ is defined as
\begin{equation}
	H_\alpha(\p) := \frac{1}{1-\alpha} \log \sum_i p_i^\alpha.
\end{equation}
To extend to quasi-distributions we must ensure the following: (a) the entropy is well-defined on quasi-distributions, and (b) obeys monotonicity under stabilizer/free operations. In the Appendices we show the following.
\begin{theorem} If $\alpha = \frac{2a}{2b-1}$ for integers $a,b \in \mathbb{N}$ and $a\ge b$ then $H_\alpha(\w)$ is well-defined on the set of quasi-distributions $\w$, and moreover if $\w \succ \w'$ for two quasi-distributions $\w, \w'$ then $H_\alpha (\w) \le H_\alpha(\w')$.
\end{theorem}
This therefore provides a meaningful extension of entropies to quasi-distributions. Moreover we can prove the following result, which establishes the equivalence between Wigner negativity and the existence of negative Renyi entropy.
\begin{theorem} A quantum state $\rho$ has Wigner negativity if and only if $H_\alpha(W_\rho) <0$ for some $\alpha =  \frac{2a}{2b-1}$, with integers $a \ge b$.
\end{theorem}
Therefore, the statistical mechanical description of stabilizer operations provides a setting in which negative entropy is fully meaningful, and quantifies the fact that magic states are `more ordered' than a perfectly sharp, deterministic classical state $\w = (1,0, \dots , 0)$ with zero entropy. We note that prior work has shown that negative conditional entropy~\cite{rio_thermodynamic_2011} arises in the context of quantum correlations, however this is unrelated to the present negative entropy.

The $\alpha \rightarrow 1$ limit of the Renyi entropy is not well-defined, however the Renyi entropy description shows that this diverges to $-\infty$ due to the Lorenz curve $L_\rho(x)$ no longer being bounded by $1$. Alternatively, we can write the divergence of the entropy as $-\epsilon^{-1} \log \sum_{\x} |W_\rho (\x)|^{1+\epsilon}$ with $\epsilon \rightarrow 0^+$, which leads to a second new description of mana: \emph{it is the rate of divergence of the Renyi-entropy} as we approach the limiting Shannon entropy $H(\w)$. We also note the recent observation~\cite{baez2011renyi} that in standard statistical mechanics that $H_\alpha(\p)$ can be related to the ``$q$--derivative'' of free energy. Therefore, we might speculate if the presence of magic in Clifford circuits admits a rigorous interpretation in terms of a phase transition signalled by a first-order divergence in the Wigner free energy. However since variation in $\alpha$ is no longer smooth ($\alpha$ is restricted to a dense set of rational values) technical obstacles exist and we do not pursue this speculative line here, but just note that other works exist in which computational transitions are considered in relation to phase transitions \ddd{[Need to cite]}.

However, every entropy is best understood as a relative entropy and so we now state the extension to Renyi divergences.
The $\alpha$--Renyi divergence $D_\alpha(\p)$ is defined as
\begin{equation}
	D_\alpha(\p|\r) := \frac{1}{\alpha-1} \log \sum_i p_i^\alpha r_i^{1-\alpha},
\end{equation}
where $\p$ is a quasi-distribution, $\r$ a probability distribution with positive components and $\alpha$ takes the same values as for $H_\alpha$.
In~\cref{app:lcst_technical}, we also show that the divergence satisfies the following natural properties for all $\alpha$ for which it is well-defined.

\begin{theorem}\label{thm:Da_props} 
	Let $\tau$ be a state in the interior of $\F$ and $\alpha = 2a / (2b-1)$ for non-negative integers $a,b$ with $a \geq b$.
Then, the $\alpha$-Renyi divergence $D_\alpha(W_\rho || W_\tau)$ is well-defined for all states $\rho$, and the following hold:
\begin{enumerate}
\item $D_\alpha(W_\rho || W_\tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\alpha(W_\rho || W_\tau) = 0$ if and only if $\rho =\tau$.
\item $D_\alpha(W_{\rho^{\otimes n}} || W_{\tau^{\otimes n}}) = n D_\alpha(W_\rho || W_\tau)$ for any $n \in \mathbb{N}$.
\item $D_\alpha(W_\rho || W_\tau) \geq D_\alpha (W_{\E(\rho)} || W_{\E(\tau)})$ for any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$.
\end{enumerate}
\end{theorem}

We can now apply this to obtain new magic bounds.
\begin{theorem}
	Consider a general magic state distillation process that converts a magic state $\rho^{\otimes n} \longrightarrow \rho'^{\otimes m}$ as well as an equilibrium state $\tau^{\otimes n} \longrightarrow \tau'^{\otimes m}$

	We have the following family of upper bounds on the distillation rate $R := n'/n$,
	\begin{equation}
		R \leq \frac{D_{\alpha}(W_\rho||W_\tau)}{D_{\alpha}(W_{\rho'}||W_{\tau'})} = \frac{F_\alpha(W_\rho) - F_\alpha(W_\tau)}{F'_\alpha(W_{\rho'}) - F'_\alpha(W_{\tau'})},
	\end{equation}
	where $\alpha = \frac{2a}{2b-1}$ for any positive integers $a,b$ with $a \geq b$ and $\beta F_\alpha(W_\rho) := D_\alpha (W_\rho || W_\tau) - \log \Z$ is the $\alpha$--free energy. 
\end{theorem}

\begin{figure}[t!]
    \centering
    \subfigure[][]{%
    \label{fig:Ra}%
    \includegraphics[scale=0.4]{figs/Ra_vs_a_eps.pdf}}\\
    \subfigure[][]{%
    \label{fig:RaDiff}%
    \includegraphics[scale=0.4]{figs/RaDiff_vs_a_eps.pdf}}
    \caption{$R_\alpha$}
    \label{fig:rate_a_contour}
\end{figure}

\section{Outlook and open questions}
\label{sec:lower_bounds}

@@

\begin{theorem} Let $\tau$ be in the interior of $\F$, and define $D_\infty(\rho,\tau) := D_\infty(W_\rho(\x) || W_\tau (\x))$. Then $D_\infty(\rho, \tau)$ is well-defined for all $\rho$, and the following hold:
\begin{enumerate}
\item $D_\infty(\rho, \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\infty(\rho , \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\infty(\rho^{\otimes 2n}, \tau^{\otimes 2n}) = n D_\infty(\rho^{\otimes 2} ,\tau^{\otimes 2})$ for any $n \in \mathbb{N}$.
\item For any quantum state $\rho$, and any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$, we have $D_\infty (\rho , \tau) \ge D_\infty (\E(\rho), \E(\tau))$.
\end{enumerate}
\end{theorem}
\begin{proof}
	Since $\tau$ is in the interior of $\F$, its Wigner function obeys $W_\tau(\x) >0$ for all points $\x$ in the phase space. In general, $W_\rho(\x)$ is a quasi-distribution, however since $\sum_{\x} W_\rho (\x) = 1$ this implies it must have at least one positive component, so $\max_{\x} \frac{W_\rho(\x)}{W_\tau(\x)}>0$. Therefore $D_\infty (\rho,\tau)$ is always well-defined.

@@

@@

We define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(W_\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(W_\rho)$ in particular is not well-defined, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(W_\rho) = 2\log_2 d = H_\alpha \left( W_{\id/d} \right),
\end{equation}
where $d$ is the dimension of $\rho$.

For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{align}
	&H_{\alpha}\left(W_{\rho_{\rm{S}}(\epsilon)}\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right] \nonumber\\
	&= \frac{\alpha}{\alpha-1}\log_2 18+ \frac{1}{1-\alpha}\log_2 \left[ 8(3-\epsilon)^\alpha + (6-8\epsilon)^\alpha \right]
\end{align}

@@

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $H_\alpha$ is decreasing in $\epsilon$ (not Schur-conocave for $\alpha < 1$) and increasing in $\alpha$.
    }
    \label{fig:Htest}
\end{figure}

@@

@@

If $\alpha =$ odd$/$odd, then $f(w)$ is not monotonous because the sign of $f'(w)$ would depend on the sign of $w$.

If $\alpha =$ even$/$odd and $\alpha < 1$, then $H_\alpha$ is Schur-convex.
}
